---
title: "Contrast RSA WebPPL Model"
author: "Dan Yurovsky and Claire Bergey"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    theme: flatly
    toc: true
    toc_float:
      collapsed: true
    code_folding: hide
---


### Setup
```{r, include=FALSE}
knitr::opts_chunk$set(fig.align='center', echo=FALSE, messages=FALSE, warning = FALSE,
                      fig.height = 3, fig.width=5)

#Load packages + setup
library(rwebppl)
library(here)
library(directlabels)
library(ggthemes)
library(lme4)
library(here)
library(ggridges)
library(tidyboot)
library(tidyverse)

theme_set(theme_few(base_size = 14))
```

# Basic contrastive inference

In this simulation, we suppose that we have two worlds: 

1. Two Dax World in which there is a red dax, a blue dax, and a blue toma
2. Two Toma world in which there is a red toma, a blue toma, a blue dax

The listener doesn't know which world they are in, but they hear a speaker say either "dax" or "blue dax" and have to pick the correct target, and also make a inference about which world they are in (and hence the meaning of "dax").

```{r two-world_inference, eval = FALSE}
two_world_utterances <- tibble(utterance = c("dax", "blue dax"),
                               utterance_num = as.character(1:2))

two_world_inference <- map_dfr(two_world_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/discrete_semantics.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(two_world_utterances, by = "utterance_num") %>%
  select(-utterance_num)
```

```{r}
contrast_plot_data <- two_world_inference %>%
  mutate(utterance = if_else(utterance == "dax", "noun", "adjective noun"),
         utterance = factor(utterance, levels = c("noun", "adjective noun")),
         obj = case_when(obj == "blue dax" & world_string == "two toma" ~ "lure",
                         obj == "blue dax" & world_string == "two dax" ~ "target",
                         obj != "blue dax" ~ NA_character_)) %>%
  filter(!is.na(obj))
  

ggplot(contrast_plot_data, aes(x = utterance, y = prob, color = obj)) + 
  geom_point()
```

```{r two-world-analysis}
two_world_obj_data <- two_world_inference %>%
  mutate(chosen = case_when(world_string == "two toma" ~ 
                           gsub("dax", "single", obj),
                         world_string == "two dax" ~ 
                           gsub("dax", "pair", obj)),
         utterance = gsub("dax", "toma", utterance),
         utterance = factor(utterance, 
                            levels = c("blue toma", "toma", "red toma")))

ggplot(two_world_obj_data, aes(x = chosen, y = prob, fill = chosen)) + 
  geom_col(position = "dodge", ) + 
  facet_wrap(~ utterance) + 
  scale_fill_ptol(drop = FALSE) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "selection probability")

ggsave("two_world_pair.png", width = 6, height = 3.5, units = "in", dpi = 600)
```

```{r markedness, eval = FALSE}
markedness_utterances <- tibble(utterance = c("toma", "red toma"),
                                utterance_num = as.character(1:2))

markedness_inference <- map_dfr(markedness_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/markedness.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(markedness_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()
```

```{r markedness-analysis}
markedness_data <- markedness_inference %>%
  mutate(utterance = gsub("red", "blue", utterance))

ggplot(markedness_data, aes(x = value, y = utterance)) + 
  stat_density_ridges(scale = 1, 
                      quantile_lines = TRUE, quantiles = 2) + 
  scale_fill_ptol() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(y = "utterance", x = "proportion of tomas that are blue")

ggsave("markedness.png", width =6, height = 4, units = "in", dpi = 600)
```

```{r combo, eval = FALSE}
contrast_utterances <- tibble(utterance = c("toma", "blue toma"),
                                utterance_num = as.character(1:2))

contrast_inference <- map_dfr(contrast_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/contrast_inference_typicality.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(contrast_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()

contrast_data <- contrast_inference %>%
  pivot_wider(names_from = "Parameter") %>%
  mutate(p = as.numeric(p))

```

```{r redundant, eval = FALSE}
redundant_utterances <- tibble(utterance = c("toma", "blue toma"),
                                utterance_num = as.character(1:2))

redundant_inference <- map_dfr(redundant_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/redundant_inference_typicality.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(redundant_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()

redundant_data <- redundant_inference %>%
  pivot_wider(names_from = "Parameter") %>%
  mutate(p = as.numeric(p)) 
```

```{r redundant, eval = FALSE}
unhelpful_utterances <- tibble(utterance = c("toma", "blue toma"),
                                utterance_num = as.character(1:2))

unhelpful_inference <- map_dfr(unhelpful_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/unhelpful_inference_typicality.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(redundant_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()

unhelpful_data <- unhelpful_inference %>%
  pivot_wider(names_from = "Parameter") %>%
  mutate(p = as.numeric(p)) 
```

```{r e2-plot}
e2_data <- contrast_data %>%
  select(-Iteration, -Chain, -world) %>%
  mutate(type = "contrast") %>%
  bind_rows(redundant_data %>% 
              select(-Iteration, -Chain, -world) %>%
              mutate(type = "redundant")) %>%
  bind_rows(unhelpful_data %>% 
              select(-Iteration, -Chain, -world) %>%
              mutate(type = "unhelpful"))

e2_data %>%
  group_by(type, utterance) %>%
  summarise(p = mean(p)) %>%
  ggplot(aes(x = type, y = p, fill = utterance)) +
  geom_col(position = "dodge") + 
  scale_fill_ptol() + 
  labs(x = "", y = "prop. tomas that are blue") + 
  theme(legend.position = "top")

ggsave("e2_rsa_nofilter.png", width = 4.5, height = 5, units = "in", dpi = 600)


e2_data %>%
  filter(obj == "blue toma") %>%
  group_by(type, utterance) %>%
  summarise(p = mean(p)) %>%
  ggplot(aes(x = type, y = p, fill = utterance)) +
  geom_col(position = "dodge") + 
  scale_fill_ptol() + 
  labs(x = "", y = "prop. tomas that are blue") + 
  theme(legend.position = "top")

ggsave("e2_rsa.png", width = 4.5, height = 5, units = "in", dpi = 600)

```


```{r joint-inference, eval = FALSE}
joint_utterances <- tibble(utterance = c("toma", "blue toma", "red toma"),
                                utterance_num = as.character(1:3))

joint_inference <- map_dfr(joint_utterances %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/two_world_typicality.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(joint_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()
```

```{r joint-analysis}
joint_data <- joint_inference %>%
  pivot_wider(names_from = "Parameter") %>%
  mutate(chosen = case_when(world == "two toma" ~ 
                           gsub("toma", "pair", obj),
                         world == "two dax" ~ 
                           gsub("toma", "single", obj))) %>%
  mutate(p = as.numeric(p),
         utterance = factor(utterance, 
                            levels = c("blue toma", "toma", "red toma")))

joint_data_obj <- joint_data %>%
  group_by(utterance, chosen) %>%
  count() %>%
  group_by(utterance) %>%
  mutate(prob = n/sum(n))

ggplot(joint_data_obj, aes(x = chosen, y = prob, fill = chosen)) + 
  geom_col(position = "dodge") + 
  facet_wrap(~ utterance) + 
  scale_fill_ptol(drop = FALSE) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "selection probability")

ggplot(joint_data, aes(x = p, y = chosen)) +
  facet_wrap(~ utterance) +
  stat_density_ridges(scale = 1, 
                      quantile_lines = TRUE, quantiles = 2) + 
  scale_fill_ptol() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(y = "object chosen", x = "proportion of tomas that are blue")

joint_data_p <- joint_data %>%
  ungroup() %>%
  filter(utterance %in% c("toma", "blue toma"), chosen == "blue pair") %>%
  select(-world, -Iteration, -Chain) %>%
  mutate(type = "contrast")


joint_data_subset <- joint_data %>%
  filter(!utterance %in% c("red dax"), !chosen == "red pair") %>%
  group_by(utterance, chosen) %>%
  summarise(p = mean(p)) %>%
  spread(utterance, p)

```

```{r reduntant-typicality, eval = FALSE}
redundant_typicality <- tibble(utterance = c("toma", "blue toma"),
                               utterance_num = as.character(1:2))

redundant_inference <- map_dfr(redundant_typicality %>% pull(utterance), 
                               ~webppl(program_file = 
                                         here("webppl/redundant_typicality.wppl"), 
                                       data = .x),
                               .id = "utterance_num") %>%
  left_join(redundant_utterances, by = "utterance_num") %>%
  select(-utterance_num) %>%
  as_tibble()
```

```{r redundant-analysis}
redundant_data <- redundant_inference %>%
  pivot_wider(names_from = "Parameter") %>%
  mutate(p = as.numeric(p)) 

ggplot(redundant_data, aes(x = p, y = world)) +
  facet_wrap(~ utterance) +
  stat_density_ridges(scale = 1, 
                      quantile_lines = TRUE, quantiles = 2) + 
  scale_fill_ptol() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(y = "object chosen", x = "proportion of tomas that are blue")

redundant_data_p <- redundant_data %>%
  filter(world %in% c("correct", "correct2")) %>%
  select(-world, -Iteration, -Chain) %>%
  mutate(type = "redundant")
```

```{r together}
comparison_data <- bind_rows(joint_data_p, redundant_data_p)

ggplot(comparison_data, aes(x = p, y = utterance)) +
  facet_wrap(~ type) +
  stat_density_ridges(scale = 1, 
                      quantile_lines = TRUE, quantiles = 2) + 
  scale_fill_ptol() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(y = "object chosen", x = "proportion of tomas that are blue")

ggsave("comparison_density.png", width = 8, height = 4, units = "in", dpi = 600)

comparison_data_avg <- comparison_data %>%
  group_by(type, utterance) %>%
  summarise(p = mean(p))

ggplot(comparison_data_avg, aes(x = utterance, y = p, color = type)) + 
  geom_point(position = "dodge")

```